---
title: "Bay Area Heat Mapping"
format: html
editor: visual
---

# Setup

## Load R Libraries

For general data wrangling that Dr. EB likes.

```{r, message = FALSE}
library(tidyverse)
library(janitor)
```

For working with map data

```{r, message = FALSE}
library(sf)
library(osmdata)
```

For imputing temperatures

```{r, message = FALSE}}
library(akima)
```

For calculating distances on the surface of the earth

```{r, message = FALSE}
library(geosphere)
```

## Load Temperature Data

### Morning

Load the \[Bay Area Head Mapping Project data\](<https://osf.io/apsx5/files/wdpb5?view_only=1cef760d3865497890a7dd51e859d8d3>) for the morning.

```{r}
am = st_read("traverses_chw_palo_alto_california_090424/palo-alto-california_am_trav.shp")
```

Plot the raw heat mapping data for morning.

```{r}
ggplot() +
  geom_sf(data = am, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red") +
  labs(colour = "Morning Temperature")
```

Load the Bay Area Head Mapping Project data for the afternoon.

### Afternoon

```{r, message = FALSE}
pm = st_read("traverses_chw_palo_alto_california_090424/palo-alto-california_pm_trav.shp")
```

## Get Impervious Surfaces Data

Read impervious surfaces data, which Anna Freitas shared with us. She said:

> I've attached the data we have on impervious surfaces. The analysis was conducted a few years ago as part of our Green Stormwater Infrastructure Plan development and did not measure impervious surface directly. Instead, staff in Watershed Protection used land use data and a standard method of estimating percent imperviousness by land use.

```{r}
impervious_surfaces = st_read("Palo_Alto_Impervious/Palo_Alto_Impervious_no open space.shp") %>% 
  select(Percent_Im)
```

Plot impervious surfaces data.

```{r}
ggplot() +
geom_sf(data = impervious_surfaces, aes(fill = Percent_Im)) +
  scale_fill_continuous(low="lightblue",high="red")
```

## Get Map Data

### Cities

Download border data for Palo Alto, East Palo Alto, and Menlo Park from \[OSM\](<https://www.openstreetmap.org/>).

(Leave commented out unless running for the first time, because this takes a long time to run)

```{r}
# borders = getbb("Palo Alto") %>%
#   opq() %>%
#   add_osm_feature(key = "boundary", value = "administrative") %>%
#   osmdata_sf()
# 
# palo_alto_border = borders$osm_multipolygons %>%
#   filter(name == "Palo Alto")
# epa_border =  borders$osm_multipolygons %>%
#   filter(name == "East Palo Alto")
# menlo_border =  borders$osm_multipolygons %>%
#   filter(name == "Menlo Park")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(menlo_border, "data/menlo_borders.gpkg")
# st_write(epa_border, "data/epa_borders.gpkg")
# st_write(palo_alto_border, "data/palo_alto_borders.gpkg")
# # Load
menlo_border = st_read("data/menlo_borders.gpkg") %>% select()
epa_border = st_read("data/epa_borders.gpkg") %>% select()
palo_alto_border = st_read("data/palo_alto_borders.gpkg") %>% select()
```

### Bay

Download SF Bay data.

```{r}
# bays = getbb("Palo Alto") %>%
#   opq() %>%
#   add_osm_feature(key = "natural", value = "bay") %>%
#   osmdata_sf()
# sf_bay = bays$osm_multipolygons %>%
#   filter(name == "San Francisco Bay")
# sf_marsh = bays$osm_multipolygons %>%
#   filter(name == "San Francisco Bay (estuarial)")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(sf_bay, "data/sf_bay.gpkg")
# st_write(sf_marsh, "data/sf_marsh.gpkg")
# Load
sf_bay = st_read("data/sf_bay.gpkg") %>% select()
sf_marsh = st_read("data/sf_marsh.gpkg") %>% select()
# # Inspect raw structure
# str(palo_alto_border)
# str(sf_bay)
```

### Coastlines

Get coastline data from OSM.

```{r}
# # getting data from coastline
# coast = osm <- getbb(place_name = "Palo Alto") %>%
#   opq() %>%
#   add_osm_feature(key = "natural", value = "coastline") %>%
#   osmdata_sf()
#  coastlines = coast$osm_lines
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(coastlines, "data/coastlines.gpkg")
# Load
coastlines = st_read("data/coastlines.gpkg") %>% select()
```

# Analysis

## Exclude Water

Graph all of SF Bay

```{r}
ggplot() +
  geom_sf(data = sf_bay)
```

Find non-water parts of Palo Alto, East Palo Alto, and Menlo Park:

```{r}
palo_alto = st_difference(palo_alto_border, sf_bay)
menlo_park = st_difference(menlo_border, sf_bay)
epa = st_difference(epa_border, sf_bay)
```

Graph city borders:

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_border, fill = "blue", alpha = 0.1)
```

Graph non-water:

```{r}
ggplot() +
  geom_sf(data = palo_alto, fill = "green", alpha = 0.1) +
  geom_sf(data = epa, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_park, fill = "blue", alpha = 0.1)
```

## Impute Temperatures

### Morning

Convert heat mapping data to a dataframe for analysis.

```{r}
tmp_data = st_coordinates(am) |>
  as.data.frame() |>
  cbind(am)
```

Impute temperatures

```{r}
granularity = 60
# Create interpolation, interp fxn associated with akima, X is lat, Y is long, Z is temp
interp_result <- with(tmp_data,
                        interp(x = X, y = Y, z = t_f,
                               xo = seq(-122.2142, -122.1057, length.out = granularity),
                               yo = seq(37.48534, 37.39437, length.out = granularity),
                              duplicate = "mean")) # handles duplicate pts

# convert this list we js made into a df for ggplot, $ pulls specific column from df
interp_df <- expand.grid(x = interp_result$x, y = interp_result$y)
interp_df$temperature <- as.vector(interp_result$z)
imputed_temperatures = interp_df %>%  drop_na()
```

Plot imputed temperatures

```{r}
# ggplot() +
#   geom_sf(data = palo_alto, fill = "green", alpha = 0.1) +
#   geom_sf(data = epa, fill = "orange", alpha = 0.1) +
#   geom_sf(data = menlo_park, fill = "blue", alpha = 0.1) +
#   geom_raster(
#     data = imputed_temperatures,
#     aes(x = x, y = y, fill=temperature),
#     alpha = 0.8
#   ) + 
#   scale_fill_continuous(low = "lightblue", high = "red") 
```
Combine cities.

```{r}
three_cities = palo_alto %>% 
  st_union(menlo_park, by_feature = TRUE) %>% 
  st_union(epa, by_feature = TRUE)
```

Convert temperature data to matching shape file.

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
am_crs = st_crs(am)
# get the coordinate system for city data
cities_crs = st_crs(three_cities)

# turn the imputed temperatures into a simple features collection (with a geometry column)
tmp = st_as_sf(imputed_temperatures, coords = c("x", "y"), crs = am_crs) |>
  # convert the coordinate system to match the projection for cities
  st_transform(cities_crs)
```

Intersect temperature data and city boundaries

```{r}
tmp_in_bounds = st_intersection(tmp, three_cities)
```

Plot where there's data.

```{r}
tmp_bbox = tmp_in_bounds %>% st_bbox()
palo_alto_with_data = st_crop(palo_alto, tmp_bbox)
epa_with_data = st_crop(epa, tmp_bbox)
menlo_park_with_data = st_crop(menlo_park, tmp_bbox)

ggplot() +
  geom_sf(data = palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_sf(
      data = tmp_in_bounds,
      aes(colour=temperature),
      alpha = 0.5
  ) +
  scale_colour_continuous(low = "lightblue", high = "red") 
```

Converting temperatures to dataframe for better plot:

```{r}
tmp_dataframe = cbind(
  st_drop_geometry(tmp_in_bounds),
  st_coordinates(tmp_in_bounds)
)
```

Slightly nicer plot of imputed temperatures within city boundaries:

```{r}
ggplot() +
  geom_sf(data = palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = tmp_dataframe,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red")
```

### Afternoon

## Linear Regression: Morning Temperature ~ Impervious Surface

Combine temperatures (within city boundaries) and impervious surface data.

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
tmp_crs = st_crs(tmp_in_bounds)
# get the coordinate system for impervious surfaces data
imp_crs = st_crs(impervious_surfaces)

# turn the imputed temperatures back into a simple features collection (with a geometry column)
tmp = tmp_in_bounds |>
  # convert the coordinate system to match the projection for impervious surfaces
  st_transform(imp_crs)

# combine the data!!!!!!
df = st_join(tmp, impervious_surfaces, join = st_within) |> clean_names()

# graph the relationship between temperature and percent_im
df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point(alpha = 0.5) # scatterplot

# do a linear regression
m = lm(temperature ~ percent_im, data = df) |>
  # compute the p-value, etc.
  summary()
m
```

Calculate margin of error for slopes

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = m$coefficients["percent_im", "Estimate"] #0.6758 
se = m$coefficients["percent_im", "Std. Error"] #0.1765
tstar = qt(p = 0.975, df = m$df[2])  #443)
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

## Linear Regression: Morning Temperature ~ Impervious Surface + Distance to Coastline

Plot impervious surface data

```{r}
ggplot() +
  geom_sf(data = palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_park_with_data, fill = "blue", alpha = 0.1) +
geom_sf(data = impervious_surfaces, aes(fill = Percent_Im), color = NA) +
  scale_fill_continuous(low="lightblue",high="red")
```

Calculate distances to coastline.

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
orig_coord_system = st_crs(df)

df_for_distances = st_transform(df, crs = "WGS84")

dist = geosphere::dist2Line(
  p = st_coordinates(df_for_distances),
  line = st_coordinates(coastlines)[,1:2]
)
```

great - now we find distance to coastline and express that via lines (actual distance is represented in feet)

```{r}
df_data = st_coordinates(df_for_distances) |>
  as.data.frame() |>
  cbind(df_for_distances)

#combine initial data with distance to coastline
df_dists <- df_data |>
  rename(y=Y,x=X) |>
  cbind(dist) |>
  mutate(miles=distance/1609)

#plot
ggplot() +
  geom_sf(data=coastlines) +
  geom_sf(data=df_for_distances, size = .1) +
  geom_segment(data=df_dists,aes(x=x,y=y,xend=lon,yend=lat), alpha = 0.1)
ggsave("pretty_star.jpg")
```

```{r}
# from dplyer
df_dists %>% 
  dplyr::select(temperature, percent_im, miles) %>% 
  drop_na()
```

```{r}
#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles, data = df_dists) |>
  # compute the p-value, etc.
  summary()
m
```

Calculate margin of error for Percent Impervious slope:

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = m$coefficients["percent_im", "Estimate"]
se = m$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = m$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

Calculate margin of error for miles from coastline slope:

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = m$coefficients["miles", "Estimate"]
se = m$coefficients["miles", "Std. Error"]
tstar = qt(p = 0.975, df = m$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

```{r}
# # margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
# point_estimate = 0.9107
# se = 0.1920
# tstar = qt(p = 0.975, df = 442)
# moe = tstar*se
# point_estimate - moe
# point_estimate + moe
# 
# # margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
# point_estimate = -0.1951
# se = 0.0657
# tstar = qt(p = 0.975, df = 442)
# moe = tstar*se
# point_estimate - moe
# point_estimate + moe
```

## Linear Regression: Afternoon Temperature ~ Impervious Surface

(repeat above, but for afternoon)

## Linear Regression: Afternoon Temperature ~ Impervious Surface + Distance to Coastline

(repeat above, but for afternoon)

## Linear Regression: Temperature ~ Impervious Surface + Distance to Coastline + Time of Day

combine datasets, with a new variable time of day that's either "morning" or "afternoon"





```{r}
st_interpolate_aw()
st_filter()
st_make_grid()
```