---
title: "Bay Area Heat Mapping"
format: html
editor: visual
---

# Setup

## Load R Libraries

For general data wrangling that Dr. EB likes.

```{r message=FALSE}
library(tidyverse)
library(janitor)
```

For working with map data

```{r message=FALSE}
library(sf)
library(osmdata)
```

For imputing temperatures

```{r message=FALSE}
library(akima)
```

For calculating distances on the surface of the earth

```{r message = FALSE}
library(geosphere)
```

## Load Temperature Data

### Morning

Load the \[Bay Area Head Mapping Project data\](<https://osf.io/apsx5/files/wdpb5?view_only=1cef760d3865497890a7dd51e859d8d3>) for the morning.

```{r}
am = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_am_trav.shp")
```

Plot the raw heat mapping data for morning.

```{r}
ggplot() +
geom_sf(data = am, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Morning)",
          color = "Morning Temperature (°F)")
```

### Afternoon

```{r}
af = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_af_trav.shp")
```

```{r}
ggplot() +
geom_sf(data = af, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Afternoon)",
          color = "Afternoon Temperature (°F)")
```

### Evening

```{r}
pm = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_pm_trav.shp")
```

```{r}
ggplot() +
geom_sf(data = pm, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Evening)",
          color = "Evening Temperature (°F)")
```

## Load Impervious surfaces data

Read impervious surfaces data, which Anna Freitas shared with us. She said:

> I've attached the data we have on impervious surfaces. The analysis was conducted a few years ago as part of our Green Stormwater Infrastructure Plan development and did not measure impervious surface directly. Instead, staff in Watershed Protection used land use data and a standard method of estimating percent imperviousness by land use.

```{r}
impervious_surfaces = st_read("data/Palo_Alto_Impervious/Palo_Alto_Impervious_no open space.shp") %>% 
  dplyr::select(Percent_Im)
```

Plot impervious surfaces data

```{r}
ggplot() +
geom_sf(data = impervious_surfaces, aes(fill = Percent_Im)) +
  scale_fill_continuous(low="lightblue",high="red")
```

## Load Map Data

### Cities

Download border data for Palo Alto, East Palo Alto, and Menlo Park from \[OSM\](<https://www.openstreetmap.org/>).

```{r}
# borders = getbb("Palo Alto") %>%
#   opq() %>%
#   add_osm_feature(key = "boundary", value = "administrative") %>%
#   osmdata_sf()
# 
# palo_alto_border = borders$osm_multipolygons %>%
#   filter(name == "Palo Alto")
# epa_border =  borders$osm_multipolygons %>%
#   filter(name == "East Palo Alto")
# menlo_border =  borders$osm_multipolygons %>%
#   filter(name == "Menlo Park")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(menlo_border, "data/menlo_borders.gpkg")
# st_write(epa_border, "data/epa_borders.gpkg")
# st_write(palo_alto_border, "data/palo_alto_borders.gpkg")
# # Load
menlo_border = st_read("data/menlo_borders.gpkg") %>% dplyr::select()
epa_border = st_read("data/epa_borders.gpkg") %>% dplyr::select()
palo_alto_border = st_read("data/palo_alto_borders.gpkg") %>% dplyr::select()
# # Inspect raw structure
# str(palo_alto_border)
# str(sf_bay)
```

### Water

```{r}
# bays = getbb("Palo Alto") %>% 
#   opq() %>% 
#   add_osm_feature(key = "natural", value = "bay") %>% 
#   osmdata_sf()
# sf_bay = bays$osm_multipolygons %>% 
#   filter(name == "San Francisco Bay")
# sf_marsh = bays$osm_multipolygons %>% 
#   filter(name == "San Francisco Bay (estuarial)")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(sf_bay, "data/sf_bay.gpkg")
# st_write(sf_marsh, "data/sf_marsh.gpkg")
# # Load
sf_bay = st_read("data/sf_bay.gpkg") %>% dplyr::select()
sf_marsh = st_read("data/sf_marsh.gpkg") %>% dplyr::select()
# # Inspect raw structure
# str(palo_alto_border)
# str(sf_bay)
```

### Coastlines

```{r}
# #getting data from coastline
# coast = osm <- getbb(place_name = "Palo Alto") %>% 
#   opq() %>% 
#   add_osm_feature(key = "natural", value = "coastline") %>% 
#   osmdata_sf()
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(coastlines, "data/coastlines.gpkg")
# Load
coastlines = st_read("data/coastlines.gpkg") %>% dplyr::select()
```

# Analysis

## Impute temperatures

### Morning

Convert temperature data to a dataframe

```{r}
am_tmp_data = cbind(
  am %>% st_drop_geometry,
  am %>% st_coordinates
)
```

Interpolate between points

```{r}
granularity = 60
minx = min(am_tmp_data$X)
maxx = max(am_tmp_data$X)
miny = min(am_tmp_data$Y)
maxy = max(am_tmp_data$Y)
# Create interpolation, interp fxn associated with akima, X is lat, Y is long, Z is temp
am_interp_result <- with(am_tmp_data,
                        interp(x = X, y = Y, z = t_f,
                               xo = seq(minx, maxx, length.out = granularity),
                               yo = seq(miny, maxy, length.out = granularity),
                              duplicate = "mean")) # handles duplicate pts

# convert this list we js made into a df for ggplot, $ pulls specific column from df
am_interp_df <- expand.grid(x = am_interp_result$x, y = am_interp_result$y)
am_interp_df$temperature <- as.vector(am_interp_result$z)
am_imputed_temperatures = am_interp_df %>%  drop_na()
```

Plot imputed temperatures

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
  # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  # geom_sf(data = menlo_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = am_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

### Afternoon


```{r}
# NOW WE FIND IMPUTED TEMPERATURES FOR AFTERNOON
af_tmp_data = st_coordinates(af) |>
  as.data.frame() |>
  cbind(af)
```

```{r}
# Create interpolation, interp fxn associated with akima, X is lat, Y is long, Z is temp
af_interp_result <- with(af_tmp_data,
                        interp(x = X, y = Y, z = t_f,
                              duplicate = "mean")) # handles duplicate pts
```

```{r}
# convert this list we js made into a df for ggplot, $ pulls specific column from df
af_interp_df <- expand.grid(x = af_interp_result$x, y =af_interp_result$y)
af_interp_df$temperature <- as.vector(af_interp_result$z)
af_imputed_temperatures = af_interp_df %>%  drop_na()
```

```{r}
af_sf_object <- st_as_sf(af_interp_df, coords = c("y", "x"), crs = 4326)
```

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
 # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
 # geom_sf(data = menlo_park_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = af_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

### Evening


```{r}
# NOW WE FIND IMPUTED TEMPERATURES FOR AFTERNOON
pm_tmp_data = st_coordinates(pm) |>
  as.data.frame() |>
  cbind(pm)
```

```{r}
# Create interpolation, interp fxn associated with akima, X is lat, Y is long, Z is temp
pm_interp_result <- with(pm_tmp_data,
                        interp(x = X, y = Y, z = t_f,
                              duplicate = "mean")) # handles duplicate pts
```

```{r}
# convert this list we js made into a df for ggplot, $ pulls specific column from df
pm_interp_df <- expand.grid(x = pm_interp_result$x, y =pm_interp_result$y)
pm_interp_df$temperature <- as.vector(pm_interp_result$z)
pm_imputed_temperatures = pm_interp_df %>%  drop_na()
```

```{r}
pm_sf_object <- st_as_sf(pm_interp_df, coords = c("y", "x"), crs = 4326)
```

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
 # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
 # geom_sf(data = menlo_park_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = pm_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

## Remove water from cities

Plot entire SF Bay

```{r}
ggplot() +
  geom_sf(data = sf_bay)
```

Exclude parts of Palo Alto, Menlo Park, and East Palo Alto that are under water.

```{r}
palo_alto = st_difference(palo_alto_border, sf_bay)
menlo_park = st_difference(menlo_border, sf_bay)
epa = st_difference(epa_border, sf_bay)
```

Plot the old city boundaries (land and water).

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_border, fill = "blue", alpha = 0.1)
```

Plot the new city boundaries (only land).

```{r}
ggplot() +
  geom_sf(data = palo_alto, fill = "green", alpha = 0.1) +
  geom_sf(data = epa, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_park, fill = "blue", alpha = 0.1)
```


## Combine Impervious Surfaces and Temperature Data

Combine temperatures (within city boundaries) and impervious surface data.

### Morning

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
am_crs = st_crs(am)

# turn the imputed temperatures into a simple features collection (with a geometry column)
am_tmp = st_as_sf(am_imputed_temperatures, coords = c("x", "y"), crs = am_crs)

sf_use_s2(FALSE)
# combine the data!!!!!!
am_combo =
  st_join(
    am_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(am_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
am_combo_df = cbind(
  st_drop_geometry(am_combo),
  st_coordinates(am_combo)
)
```

Crop city boundaries to where there's data

```{r}
am_combo_tmp_bbox = am_combo %>% st_bbox()
am_palo_alto_with_data = st_crop(palo_alto, am_combo_tmp_bbox)
am_epa_with_data = st_crop(epa, am_combo_tmp_bbox)
am_menlo_park_with_data = st_crop(menlo_park, am_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap

```{r}
ggplot() +
  geom_sf(data = am_palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = am_epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = am_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = am_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Imputed Temperatures (Morning)",
    fill = "Morning Temperature (°F)")
ggsave("am_temperatures.png")
```

### Afternoon

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
af_crs = st_crs(af)

# turn the imputed temperatures into a simple features collection (with a geometry column)
af_tmp = st_as_sf(af_imputed_temperatures, coords = c("x", "y"), crs = af_crs)

# combine the data!!!!!!
af_combo =
  st_join(
    af_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(af_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
af_combo_df = cbind(
  st_drop_geometry(af_combo),
  st_coordinates(af_combo)
)
```

Crop city boundaries to where there's data

```{r}
af_combo_tmp_bbox = af_combo %>% st_bbox()
af_palo_alto_with_data = st_crop(palo_alto, af_combo_tmp_bbox)
af_epa_with_data = st_crop(epa, af_combo_tmp_bbox)
af_menlo_park_with_data = st_crop(menlo_park, af_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap

```{r}
ggplot() +
  geom_sf(data = af_palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = af_epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = af_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = af_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Imputed Temperatures (Afternoon)",
    fill = "Afternoon Temperature (°F)")
ggsave("af_temperatures.png")
```

### Evening

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
pm_crs = st_crs(pm)

# turn the imputed temperatures into a simple features collection (with a geometry column)
pm_tmp = st_as_sf(pm_imputed_temperatures, coords = c("x", "y"), crs = pm_crs)

# combine the data!!!!!!
pm_combo =
  st_join(
    pm_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(pm_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
pm_combo_df = cbind(
  st_drop_geometry(pm_combo),
  st_coordinates(pm_combo)
)
```

Crop city boundaries to where there's data

```{r}
pm_combo_tmp_bbox = pm_combo %>% st_bbox()
pm_palo_alto_with_data = st_crop(palo_alto, pm_combo_tmp_bbox)
pm_epa_with_data = st_crop(epa, pm_combo_tmp_bbox)
pm_menlo_park_with_data = st_crop(menlo_park, pm_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap

```{r}
ggplot() +
  geom_sf(data = pm_palo_alto_with_data, fill = "green", alpha = 0.1) +
  geom_sf(data = pm_epa_with_data, fill = "orange", alpha = 0.1) +
  geom_sf(data = pm_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = pm_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Imputed Temperatures (Evening)",
    fill = "Evening Temperature (°F)")
ggsave("pm_temperatures.png")
```
## Calculate distances to coastline.

### Morning

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
am_orig_coord_system = st_crs(am_combo)

am_combo_for_distances = st_transform(am_combo, crs = "WGS84")
# 
# am_df_data = st_coordinates(am_combo_for_distances) |>
#   as.data.frame() |>
#   cbind(am_combo_for_distances)
# 
# am_dist = geosphere::dist2Line(
#   p = st_coordinates(am_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# am_df_dists <- am_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(am_df_dists, "data/am_df_dists.csv")
# Load
am_df_dists = read_csv("data/am_df_dists.csv")
```

great - now we find distance to coastline and express that via lines (actual distance is represented in feet)

```{r}
#plot
ggplot() +
  geom_sf(data=coastlines) +
  geom_sf(data=am_combo_for_distances, size = .1) +
  geom_segment(data=am_df_dists,aes(x=x,y=y,xend=lon,yend=lat), alpha = 0.1)
ggsave("pretty_star.jpg")
```


### Afternoon

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
af_orig_coord_system = st_crs(af_combo)

af_combo_for_distances = st_transform(af_combo, crs = "WGS84")

af_df_data = st_coordinates(af_combo_for_distances) |>
  as.data.frame() |>
  cbind(af_combo_for_distances)

# af_dist = geosphere::dist2Line(
#   p = st_coordinates(af_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# af_df_dists <- af_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(af_dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(af_df_dists, "data/af_df_dists.csv")
# Load
af_df_dists = read_csv("data/af_df_dists.csv")
```

### Evening

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
pm_orig_coord_system = st_crs(pm_combo)

pm_combo_for_distances = st_transform(pm_combo, crs = "WGS84")

pm_df_data = st_coordinates(pm_combo_for_distances) |>
  as.data.frame() |>
  cbind(pm_combo_for_distances)

# pm_dist = geosphere::dist2Line(
#   p = st_coordinates(pm_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# pm_df_dists <- pm_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(pm_dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(pm_df_dists, "data/pm_df_dists.csv")
# Load
pm_df_dists = read_csv("data/pm_df_dists.csv")
```

## Linear Regressions

### Morning Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
am_combo_df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

# do a linear regression
am_lm_results = lm(temperature ~ percent_im, data = am_combo_df) |>
  # compute the p-value, etc.
  summary()
am_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = am_lm_results$coefficients["percent_im", "Estimate"]
se = am_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = am_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

### Afternoon Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
af_combo_df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

# do a linear regression
af_lm_results = lm(temperature ~ percent_im, data = af_combo_df) |>
  # compute the p-value, etc.
  summary()
af_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = af_lm_results$coefficients["percent_im", "Estimate"]
se = af_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = af_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

### Evening Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
pm_combo_df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

# do a linear regression
pm_lm_results = lm(temperature ~ percent_im, data = pm_combo_df) |>
  # compute the p-value, etc.
  summary()
pm_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = pm_lm_results$coefficients["percent_im", "Estimate"]
se = pm_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = pm_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```


### Morning Temperature \~ Impervious Surface + Distance to Coastline

```{r}
#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles, data = df_dists) |>
  # compute the p-value, etc.
  summary()
m
```

Calculate margin of error for Percent Impervious slope:

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = m$coefficients["percent_im", "Estimate"]
se = m$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = m$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

Calculate margin of error for miles from coastline slope:

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = m$coefficients["miles", "Estimate"]
se = m$coefficients["miles", "Std. Error"]
tstar = qt(p = 0.975, df = m$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```
