---
title: "Palo Alto Heat Mapping"
format: html
editor: visual
---

# Setup

## Load R Libraries

```{r message=FALSE}
install.packages("broom")
install.packages("modelr")
```

```{r message=FALSE}
library(broom)
library(modelr)
```

```{r}
install.packages("tidyverse")
```

```{r message = FALSE}
install.packages("janitor")
```

For general data wrangling that Dr. EB likes.

```{r}
library(tidyverse)
library(janitor)
```

```{r message=FALSE}
library(modelr)
```

```{r}
install.packages("sf")
```

```{r}
install.packages("osmdata")
```

For working with map data

```{r message=FALSE}
library(sf)
library(osmdata)
```

```{r}
install.packages("akima")
```

For imputing temperatures

```{r message=FALSE}
library(akima)
```

For calculating distances on the surface of the earth

```{r}
install.packages("geosphere"
)
```

```{r message = FALSE}
library(geosphere)
```

## Load Temperature Data

### Morning

Load the \[Bay Area Head Mapping Project data\](<https://osf.io/apsx5/files/wdpb5?view_only=1cef760d3865497890a7dd51e859d8d3>) for the morning.

```{r}
am = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_am_trav.shp")
```
```{r}
# 1. Read the shapefile as an sf object
am_raw_df <- st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_am_trav.shp")

# 2. Convert to a standard data frame by dropping geometry
am_traversal_df <- st_drop_geometry(am_raw_df)
```
```{r}
#find morning (traversal) average temperature, temperature range
fivenum(am_traversal_df %>% 
       pull(t_f))
mean(am_traversal_df %>% 
       pull(t_f))
```



Plot the raw heat mapping data for morning.

```{r}
ggplot() +
geom_sf(data = am, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Morning 6-7 AM)",
          color = "Morning Temperature (°F)")
```
### Afternoon

```{r}
af = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_af_trav.shp")
```
```{r}
# 1. Read the shapefile as an sf object
af_raw_df <- st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_af_trav.shp")

# 2. Convert to a standard data frame by dropping geometry
af_traversal_df <- st_drop_geometry(af_raw_df)

#find morning (traversal) average temperature, temperature range
fivenum(af_traversal_df %>% 
       pull(t_f))
mean(af_traversal_df %>% 
       pull(t_f))
```

```{r}
ggplot() +
geom_sf(data = af, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Afternoon 3-4 PM)",
          color = "Afternoon Temperature (°F)")
```

### Evening

```{r}
pm = st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_pm_trav.shp")
```
```{r}
# 1. Read the shapefile as an sf object
pm_raw_df <- st_read("data/traverses_chw_palo_alto_california_090424/palo-alto-california_pm_trav.shp")

# 2. Convert to a standard data frame by dropping geometry
pm_traversal_df <- st_drop_geometry(pm_raw_df)

#find morning (traversal) average temperature, temperature range
fivenum(pm_traversal_df %>% 
       pull(t_f))
mean(pm_traversal_df %>% 
       pull(t_f))
```

```{r}
ggplot() +
geom_sf(data = pm, aes(color = t_f)) +
  scale_color_continuous(low="lightblue",high="red"
                         ) + 
  labs(title = "Continuous Traversal Points (Evening 7-8 PM)",
          color = "Evening Temperature (°F)")
```

## Load Impervious surfaces data

Read impervious surfaces data, which Anna Freitas shared with us. She said:

> I've attached the data we have on impervious surfaces. The analysis was conducted a few years ago as part of our Green Stormwater Infrastructure Plan development and did not measure impervious surface directly. Instead, staff in Watershed Protection used land use data and a standard method of estimating percent imperviousness by land use.

```{r}
impervious_surfaces = st_read("data/Palo_Alto_Impervious/Palo_Alto_Impervious_no open space.shp") %>% 
  dplyr::select(Percent_Im)
```
```{r}
# 1. Read the shapefile as an sf object
imperv_surf_df <- st_read("data/Palo_Alto_Impervious/Palo_Alto_Impervious_no open space.shp")

# 2. Convert to a standard data frame by dropping geometry
imperv_surface_df <- st_drop_geometry(imperv_surf_df)

#find morning (traversal) average temperature, temperature range
fivenum(imperv_surface_df %>% 
       pull(Percent_Im))
mean(imperv_surface_df %>% 
       pull(Percent_Im))
```

Plot impervious surfaces data

```{r}
ggplot() +
geom_sf(data = impervious_surfaces, aes(fill = Percent_Im)) +
  scale_fill_continuous(low="lightblue",high="red") +
  labs(title = "Palo Alto Impervious Surface",
       fill = "Percent Impervious Surface") 
```

## Load Map Data

### Cities

Download border data for Palo Alto, East Palo Alto, and Menlo Park from \[OSM\](<https://www.openstreetmap.org/>).

```{r}
# borders = getbb("Palo Alto") %>%
#   opq() %>%
#   add_osm_feature(key = "boundary", value = "administrative") %>%
#   osmdata_sf()
# 
# palo_alto_border = borders$osm_multipolygons %>%
#   filter(name == "Palo Alto")
# epa_border =  borders$osm_multipolygons %>%
#   filter(name == "East Palo Alto")
# menlo_border =  borders$osm_multipolygons %>%
#   filter(name == "Menlo Park")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(menlo_border, "data/menlo_borders.gpkg")
# st_write(epa_border, "data/epa_borders.gpkg")
# st_write(palo_alto_border, "data/palo_alto_borders.gpkg")
# # Load
menlo_border = st_read("data/menlo_borders.gpkg") %>% dplyr::select()
epa_border = st_read("data/epa_borders.gpkg") %>% dplyr::select()
palo_alto_border = st_read("data/palo_alto_borders.gpkg") %>% dplyr::select()
# # Inspect raw structure
# str(palo_alto_border)
# str(sf_bay)
```

### Water

```{r}
# bays = getbb("Palo Alto") %>% 
#   opq() %>% 
#   add_osm_feature(key = "natural", value = "bay") %>% 
#   osmdata_sf()
# sf_bay = bays$osm_multipolygons %>% 
#   filter(name == "San Francisco Bay")
# sf_marsh = bays$osm_multipolygons %>% 
#   filter(name == "San Francisco Bay (estuarial)")
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(sf_bay, "data/sf_bay.gpkg")
# st_write(sf_marsh, "data/sf_marsh.gpkg")
# # Load
sf_bay = st_read("data/sf_bay.gpkg") %>% dplyr::select()
sf_marsh = st_read("data/sf_marsh.gpkg") %>% dplyr::select()
# # Inspect raw structure
# str(palo_alto_border)
# str(sf_bay)
```

### Coastlines

```{r}
# #getting data from coastline
# coast = osm <- getbb(place_name = "Palo Alto") %>% 
#   opq() %>% 
#   add_osm_feature(key = "natural", value = "coastline") %>% 
#   osmdata_sf()
```

Cache/load the data locally.

```{r}
# # Cache
# st_write(coastlines, "data/coastlines.gpkg")
# Load
coastlines = st_read("data/coastlines.gpkg") %>% dplyr::select()
```

# Analysis

## Impute temperatures

Make a function to impute temperatures for any time of day

```{r}
impute_temperatures = function(df, granularity) {
  # Convert temperature data to a dataframe
  tmp_data = cbind(
    df %>% st_drop_geometry,
    df %>% st_coordinates
  )
  
  # Interpolate between points
  minx = min(tmp_data$X)
  maxx = max(tmp_data$X)
  miny = min(tmp_data$Y)
  maxy = max(tmp_data$Y)
  # Create interpolation, interp fxn associated with akima, X is lat, Y is long, Z is temp
  interp_result <- with(tmp_data,
                           interp(x = X, y = Y, z = t_f,
                                  xo = seq(minx, maxx, length.out = granularity),
                                  yo = seq(miny, maxy, length.out = granularity),
                                  duplicate = "mean")) # handles duplicate pts
  
  # convert this list we js made into a df for ggplot, $ pulls specific column from df
  interp_df <- expand.grid(x = interp_result$x, y = interp_result$y)
  interp_df$temperature <- as.vector(interp_result$z)
  interp_df %>%  drop_na()
}
```



### Morning

```{r}
am_imputed_temperatures = impute_temperatures(am, 60)
```

Plot imputed temperatures

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
  # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  # geom_sf(data = menlo_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = am_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

### Afternoon

```{r}
af_imputed_temperatures = impute_temperatures(af, 60)
```

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
  geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  geom_sf(data = menlo_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = af_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

### Evening

```{r}
pm_imputed_temperatures = impute_temperatures(pm, 60)
```

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1) +
 # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
 # geom_sf(data = menlo_park_border, fill = "blue", alpha = 0.1) +
geom_raster(
    data = pm_imputed_temperatures,
    aes(x = x, y = y, fill=temperature),
    alpha = 0.8
) + 
  scale_fill_continuous(low = "lightblue", high = "red") 
```

## Remove water from cities

Plot entire SF Bay

```{r}
ggplot() +
  geom_sf(data = sf_bay)
```

Exclude parts of Palo Alto, Menlo Park, and East Palo Alto that are under water.

```{r}
palo_alto = st_difference(palo_alto_border, sf_bay)
#menlo_park = st_difference(menlo_border, sf_bay)
#epa = st_difference(epa_border, sf_bay)
```

Plot the old city boundaries (land and water).

```{r}
ggplot() +
  geom_sf(data = palo_alto_border, fill = "green", alpha = 0.1)
 # geom_sf(data = epa_border, fill = "orange", alpha = 0.1) +
  #geom_sf(data = menlo_border, fill = "blue", alpha = 0.1)
```

Plot the new city boundaries (only land).

```{r}
ggplot() +
  geom_sf(data = palo_alto, fill = "green", alpha = 0.1)
 # geom_sf(data = epa, fill = "orange", alpha = 0.1) +
  #geom_sf(data = menlo_park, fill = "blue", alpha = 0.1)
```

## Combine Impervious Surfaces and Temperature Data

Combine temperatures (within city boundaries) and impervious surface data.

### Morning

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
am_crs = st_crs(am)

# turn the imputed temperatures into a simple features collection (with a geometry column)
am_tmp = st_as_sf(am_imputed_temperatures, coords = c("x", "y"), crs = am_crs)

sf_use_s2(FALSE)
# combine the data!!!!!!
am_combo =
  st_join(
    am_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(am_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
am_combo_df = cbind(
  st_drop_geometry(am_combo),
  st_coordinates(am_combo)
)
```

Crop city boundaries to where there's data

```{r}
am_combo_tmp_bbox = am_combo %>% st_bbox()
am_palo_alto_with_data = st_crop(palo_alto, am_combo_tmp_bbox)
#am_epa_with_data = st_crop(epa, am_combo_tmp_bbox)
#am_menlo_park_with_data = st_crop(menlo_park, am_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap -\> only Palo Alto

```{r}
ggplot() +
  geom_sf(data = am_palo_alto_with_data, fill = "green", alpha = 0.1) +
 # geom_sf(data = am_epa_with_data, fill = "orange", alpha = 0.1) +
 # geom_sf(data = am_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = am_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Palo Alto Imputed Temperatures (Morning 6-7 AM)",
    fill = "Morning Temperature (°F)")
ggsave("am_temperatures.png")
```

### Afternoon

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
af_crs = st_crs(af)

# turn the imputed temperatures into a simple features collection (with a geometry column)
af_tmp = st_as_sf(af_imputed_temperatures, coords = c("x", "y"), crs = af_crs)

# combine the data!!!!!!
af_combo =
  st_join(
    af_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(af_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
af_combo_df = cbind(
  st_drop_geometry(af_combo),
  st_coordinates(af_combo)
)
```

Crop city boundaries to where there's data

```{r}
af_combo_tmp_bbox = af_combo %>% st_bbox()
af_palo_alto_with_data = st_crop(palo_alto, af_combo_tmp_bbox)
# af_epa_with_data = st_crop(epa, af_combo_tmp_bbox)
# af_menlo_park_with_data = st_crop(menlo_park, af_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap

```{r}
ggplot() +
  geom_sf(data = af_palo_alto_with_data, fill = "green", alpha = 0.1) +
 # geom_sf(data = af_epa_with_data, fill = "orange", alpha = 0.1) +
 # geom_sf(data = af_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = af_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Palo Alto Imputed Temperatures (Afternoon 3-4 PM)",
    fill = "Afternoon Temperature (°F)")
ggsave("af_temperatures.png")
```

### Evening

```{r}
# different data uses different projections of the globe onto a 2-d plane
# get the coordinate system for am data
pm_crs = st_crs(pm)

# turn the imputed temperatures into a simple features collection (with a geometry column)
pm_tmp = st_as_sf(pm_imputed_temperatures, coords = c("x", "y"), crs = pm_crs)

# combine the data!!!!!!
pm_combo =
  st_join(
    pm_tmp,
    # convert the coordinate system to match the projection for tempterature data
    impervious_surfaces |> st_transform(st_crs(pm_tmp)),
    join = st_within
  ) |>
  clean_names() |>
  drop_na()

# convert to dataframe for raster plot
pm_combo_df = cbind(
  st_drop_geometry(pm_combo),
  st_coordinates(pm_combo)
)
```

Crop city boundaries to where there's data

```{r}
pm_combo_tmp_bbox = pm_combo %>% st_bbox()
pm_palo_alto_with_data = st_crop(palo_alto, pm_combo_tmp_bbox)
#pm_epa_with_data = st_crop(epa, pm_combo_tmp_bbox)
#pm_menlo_park_with_data = st_crop(menlo_park, pm_combo_tmp_bbox)
```

Plot imputed temperature and surface data overlap

```{r}
ggplot() +
  geom_sf(data = pm_palo_alto_with_data, fill = "green", alpha = 0.1) +
  #geom_sf(data = pm_epa_with_data, fill = "orange", alpha = 0.1) +
  #geom_sf(data = pm_menlo_park_with_data, fill = "blue", alpha = 0.1) +
  geom_raster(
    data = pm_combo_df,
    aes(x = X, y = Y, fill=temperature),
    alpha = 0.8
  ) +
  scale_fill_continuous(low = "lightblue", high = "red") +
  labs(
    x = NULL,
    y = NULL,
    title = "Imputed Temperatures (Evening 7-8 PM)",
    fill = "Evening Temperature (°F)")
ggsave("pm_temperatures.png")
```

## Calculate distances to coastline.

### Morning

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
am_orig_coord_system = st_crs(am_combo)

am_combo_for_distances = st_transform(am_combo, crs = "WGS84")
# 
# am_df_data = st_coordinates(am_combo_for_distances) |>
#   as.data.frame() |>
#   cbind(am_combo_for_distances)
# 
# am_dist = geosphere::dist2Line(
#   p = st_coordinates(am_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# am_df_dists <- am_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(am_df_dists, "data/am_df_dists.csv")
# Load
am_df_dists = read_csv("data/am_df_dists.csv")
```

great - now we find distance to coastline and express that via lines (actual distance is represented in feet)

```{r}
#plot
ggplot() +
  geom_sf(data=coastlines) +
  geom_sf(data=am_combo_for_distances, size = .1) +
  geom_segment(data=am_df_dists,aes(x=x,y=y,xend=lon,yend=lat), alpha = 0.1)
ggsave("pretty_star.jpg")
```

### Afternoon

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
af_orig_coord_system = st_crs(af_combo)

af_combo_for_distances = st_transform(af_combo, crs = "WGS84")

af_df_data = st_coordinates(af_combo_for_distances) |>
  as.data.frame() |>
  cbind(af_combo_for_distances)

# af_dist = geosphere::dist2Line(
#   p = st_coordinates(af_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# af_df_dists <- af_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(af_dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(af_df_dists, "data/af_df_dists.csv")
# Load
af_df_dists = read_csv("data/af_df_dists.csv")
```

### Evening

```{r}
# use dist2line from geosphere - only works for WGS84 (coordinate system)
pm_orig_coord_system = st_crs(pm_combo)

pm_combo_for_distances = st_transform(pm_combo, crs = "WGS84")

pm_df_data = st_coordinates(pm_combo_for_distances) |>
  as.data.frame() |>
  cbind(pm_combo_for_distances)

# pm_dist = geosphere::dist2Line(
#   p = st_coordinates(pm_combo_for_distances),
#   line = st_coordinates(coastlines)[,1:2]
# )
# 
# #combine initial data with distance to coastline
# pm_df_dists <- pm_df_data |>
#   rename(y=Y,x=X) |>
#   cbind(pm_dist) |>
#   mutate(miles=distance/1609)
```

```{r}
# # Cache
# write_csv(pm_df_dists, "data/pm_df_dists.csv")
# Load
pm_df_dists = read_csv("data/pm_df_dists.csv")
```

#Combining The Data

```{r}
gen_temp = bind_rows(
  list(
am_df_dists %>% 
  mutate(time = "6-7am"),
af_df_dists  %>% 
  mutate(time = "3-4pm"),
pm_df_dists  %>% 
  mutate(time = "7-8pm")
)
) %>% 
  # mutate(
  #   time_from_solar_noon = case_when(
  #     time == "6-7am" ~ -.5,
  #       time == "3-4pm" ~ 3.5 
  #   )
  # )
  mutate(milessq = miles^2,
         milescubed = miles^3,
         milesfourth = miles^4,
         lnmiles = log(miles))
```

##Finding coefficients for fitting combined data

###6-7am

```{r}
gen_temp %>% 
  filter(time == "6-7am") %>% 
  lm(temperature ~ x + x^2, data = .) %>% 
  summary()
```

###6-7 am fitted graph - quadratic

```{r}
# c = 5.689*10^1
# b = 1.118*10^(-3)
# a  = -2.900*10^(-7)
# df_augmented = gen_temp %>% 
#   filter(time == "6-7am") %>% 
#   mutate(prediction = a*x^2 + b*x + c,
#          resid = temperature - prediction)
# df_augmented %>% 
#   ggplot() +
#   aes(x = distance, y = temperature, colour = time) +
#   geom_point() +
#  geom_function(fun = function(x) {a*x^2 + b*x + c},
#                color = "red", lwd = 2) +
#   # geom_smooth(method = "loess") +
#   
#   facet_wrap(vars(time),scales = "free")
```

```{r}
#finding average temperature for 6-7 am
mean(gen_temp %>% 
  filter(time == "6-7am") %>% pull(temperature))
```

```{r}
#finding average percent imp
mean(gen_temp %>% 
  filter(time == "6-7am") %>% 
     pull(percent_im))
```

###6-7 am fitted graph - log

```{r}
a = 1.02383-1.07168
b = 91 + 1.27703 -35.00931
df_augmented = gen_temp %>% 
  filter(time == "6-7am") %>% 
  mutate(prediction = a*log(x)+b,
         resid = temperature - prediction)
df_augmented %>% 
  ggplot() +
  aes(x = miles, y = temperature, colour = time) +
  geom_point() +
 geom_function(fun = function(x) {a*log(x)+b},
               color = "red", lwd = 2) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```
###6-7 am fitted graph - square root (maybe unnecessary)
```{r}
a = 1.02383-1.07168
b = 91 + 1.27703 -35.00931
df_augmented = gen_temp %>% 
  filter(time == "6-7am") %>% 
  mutate(prediction = a*sqrt(x)+b,
         resid = temperature - prediction)
df_augmented %>% 
  ggplot() +
  aes(x = miles, y = temperature, colour = time) +
  geom_point() +
 geom_function(fun = function(x) {a*sqrt(x)+b},
               color = "red", lwd = 2) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###3-4 pm - loess

```{r}
c = 8.955e+01
b = 2.466e-03
a  = -3.656e-07
gen_temp %>% 
   filter(time == "3-4pm") %>% 
  ggplot() +
  aes(x = distance, y = temperature, colour = time) +
geom_point() +
  geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###3-4 pm fitted graph - poly

```{r}
c = 8.955e+01
b = 2.466e-03
a  = -3.656e-07
gen_temp %>% 
   filter(time == "3-4pm") %>% 
  ggplot() +
  aes(x = distance, y = temperature, colour = time) +
geom_point() +
  geom_function(fun = function(x) {a*x^2 + b*x + c}) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###3-4 pm fitted graph - loess

```{r}
c = 8.955e+01
b = 2.466e-03
a  = -3.656e-07
gen_temp %>% 
   filter(time == "3-4pm") %>% 
  ggplot() +
  aes(x = distance, y = temperature, colour = time) +
geom_point() +
 # geom_function(fun = function(x) {a*x^2 + b*x + c}) +
   geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###3-4 pm fitted graph - log

```{r}
a = 1.02383 
b = 91.83215 + 0.5597818*1.27703 
df_augmented = gen_temp %>% 
  filter(time == "3-4pm") %>% 
  mutate(prediction = a*log(miles)+b,
         resid = temperature - prediction)
df_augmented %>% 
  ggplot() +
  aes(x = miles, y = temperature, colour = time) +
  geom_point() +
 geom_function(fun = function(x) {a*log(x)+b},
               color = "red", lwd = 2) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###7-8 pm

```{r}
gen_temp %>% 
  filter(time == "7-8pm") %>% 
  lm(temperature ~ x + x^2, data = .) %>% 
  summary()
```

###7-8 pm fitted graph - poly

```{r}
c = 7.720e+01
b = 4.806e-03
a  = -6.596e-07
gen_temp %>% 
   filter(time == "7-8pm") %>% 
  ggplot() +
  aes(x = distance, y = temperature, colour = time) +
geom_point() +
  geom_function(fun = function(x) {a*x^2 + b*x + c}) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###7-8 pm fitted graph - loess

```{r}
c = 7.720e+01
b = 4.806e-03
a  = -6.596e-07
gen_temp %>% 
   filter(time == "7-8pm") %>% 
  ggplot() +
  aes(x = distance, y = temperature, colour = time) +
geom_point() +
 # geom_function(fun = function(x) {a*x^2 + b*x + c}) +
   geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

###7-8 pm fitted graph - log

```{r}
#why is my y-axis off?
a = 1.02383 + 1.10388
b = 91 + 0.5597818*1.27703 - 9.23545 
df_augmented = gen_temp %>% 
  filter(time == "7-8pm") %>% 
  mutate(prediction = a*log(x)+b,
         resid = temperature - prediction)
df_augmented %>% 
  ggplot() +
  aes(x = miles, y = temperature, colour = time) +
  geom_point() +
 geom_function(fun = function(x) {a*log(x)+b},
               color = "red", lwd = 2) +
  # geom_smooth(method = "loess") +
  
  facet_wrap(vars(time),scales = "free")
```

##Finding Residuals of Fitted Graphs Using Impervious Surface Data ###6-7am residuals

```{r}
# Plot residuals against impervious surface ('impervious_surfaces') to check for patterns - it has a prob w/ .resid (i think i need to name it tbh), okay so i named it -> has a problem with length: slight issue since impervious surface data is by land usage not the same places as .resid (can i bypass this somehow?); can't seem to bypass :(
ggplot(df_augmented, aes(x = percent_im, y = resid)) +
  geom_point() +
  geom_smooth(method = "loess", color = "red") + # Smoother to highlight trend
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Impervious Surfaces",
       x = "Impervous Surfaces (%)",
       y = "Residuals")
```

```{r}
lm(temperature ~ distance * percent_im * time, data = gen_temp) %>% 
  summary()
```

```{r}
gen_temp %>% 
  ggplot() +
  aes(x = percent_im, y = temperature, colour = time) +
geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(vars(time),scales = "free") +
  labs(x = "Impervious Surface Percentage",
       y = "Air Temperature (°F)",
       title = "Air Temperature (°F) vs. Impervious Surface Percentage")
```

## Linear Regressions - shouldn't be linear (acknowledge this in paper/poster)
-> figure out why it isn't linear
### Morning Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
am_combo_df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm",
              color = "black") + 
  geom_point(color = "red") +
  labs(x = "Percent Impervious Surface",
       y = "Morning Air Temperature (°F)",
       title = "Morning Air Temperature (°F) vs. Percent Impervious Surface")

# do a linear regression
am_lm_results = lm(temperature ~ percent_im, data = am_combo_df) |>
  # compute the p-value, etc.
  summary()
am_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = am_lm_results$coefficients["percent_im", "Estimate"]
se = am_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = am_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

### Afternoon Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
af_combo_df |>
    ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm",
              color = "black") + 
  geom_point(color = "red") +
  labs(x = "Percent Impervious Surface",
       y = "Afternoon Air Temperature (°F)",
       title = "Afternoon Air Temperature (°F) vs. Percent Impervious Surface")

# do a linear regression
af_lm_results = lm(temperature ~ percent_im, data = af_combo_df) |>
  # compute the p-value, etc.
  summary()
af_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = af_lm_results$coefficients["percent_im", "Estimate"]
se = af_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = af_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

### Evening Temperature \~ Impervious Surface

```{r}
# graph the relationship between temperature and percent_im
pm_combo_df |>
  ggplot() +
  aes(x = percent_im, y = temperature) +
  geom_smooth(method = "lm",
              color = "black") + 
  geom_point(color = "red") +
  labs(x = "Percent Impervious Surface",
       y = "Evening Air Temperature (°F)",
       title = "Evening Air Temperature (°F) vs. Percent Impervious Surface")

# do a linear regression
pm_lm_results = lm(temperature ~ percent_im, data = pm_combo_df) |>
  # compute the p-value, etc.
  summary()
pm_lm_results
```

Calculate margin of error for slopes (automatically read from the model summary `m`!)

```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT
point_estimate = pm_lm_results$coefficients["percent_im", "Estimate"]
se = pm_lm_results$coefficients["percent_im", "Std. Error"]
tstar = qt(p = 0.975, df = pm_lm_results$df[2])
moe = tstar*se
point_estimate - moe
point_estimate + moe
```

### Temperature \~ Impervious Surface + Distance to Coastline(quadratic) + Time

```{r}
#mid day (3-4 pm) is baseline; everything compared it; (Intercept) is avg temp @ 3-4 pm
#this is a bad graph - don't include
gen_temp |>
  ggplot() +
  aes(x = miles, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() +
  labs(title = "Bad Graph")

#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles + milessq + time, data = gen_temp) 
m |>
  # compute the p-value, etc.
  summary()

df_augmented =
  gen_temp %>% 
  mutate(
    predicted = predict(m),
    residual=resid(m)
  )
```

```{r}
# Plot residuals against impervious surface ('impervious_surfaces') to check for patterns - it has a prob w/ .resid (i think i need to name it tbh), okay so i named it -> has a problem with length: slight issue since impervious surface data is by land usage not the same places as .resid (can i bypass this somehow?); can't seem to bypass :(
ggplot(df_augmented %>% filter(time == "6-7am"), aes(x = predicted, y = residual)) +
  geom_point() +
  geom_smooth(method = "loess", color = "red") + # Smoother to highlight trend
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "6-7 am Residuals vs Impervious Surfaces",
       x = "Impervous Surfaces (%)",
       y = "Residuals")
```

### Temperature \~ Impervious Surface + Distance to Coastline(log) + Time

######IMPORTANT NUMBERS + SHIT

```{r}
#mid day (3-4 pm) is baseline; everything compared it; (Intercept) is avg temp @ 3-4 pm
gen_temp |>
  ggplot() +
  aes(x = miles, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + lnmiles * time, data = gen_temp) 
m |>
  # compute the p-value, etc.
  summary()

df_augmented =
  gen_temp %>% 
  mutate(
    predicted = predict(m),
    residual=resid(m)
  )
```

```{r}
# Plot residuals against impervious surface ('impervious_surfaces') to check for patterns - it has a prob w/ .resid (i think i need to name it tbh), okay so i named it -> has a problem with length: slight issue since impervious surface data is by land usage not the same places as .resid (can i bypass this somehow?); can't seem to bypass :(
ggplot(df_augmented %>% filter(time == "6-7am"), aes(x = predicted, y = residual)) +
  geom_point() +
  geom_smooth(method = "loess", color = "red") + # Smoother to highlight trend
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "6-7 am Residuals vs Impervious Surfaces",
       x = "Impervous Surfaces (%)",
       y = "Residuals")
```
```{r}
m = lm(
  temperature ~ 
    percent_im +
    time +
    time:miles +
    time:milessq,
  data = gen_temp
)
m %>% summary()

coef = (m %>% summary())$coefficient[, "Estimate"]

str_c("temperature = ", str_c(names(coef), " * ", coef, collapse = " + "))

coef
```

Calculate margin of error for Percent Impervious slope:


```{r}
# margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT --> commenting out b/c it's bugging out on me
# point_estimate = m$coefficients["percent_im", "Estimate"]
# se = m$coefficients["percent_im", "Std. Error"]
# tstar = qt(p = 0.975, df = m$df[2])
# moe = tstar*se
# point_estimate - moe
# point_estimate + moe
```



Calculate margin of error for miles from coastline slope:

Manually calculating the margin of error:
point_estimate for slope of miles vs estimated temperature = 

```{r}
# # margin of error for slopes!!! doesn't include 0 SOOO WE ARE STATISTICALLY SIGNIFICANT --> commenting out b/c it throws a bug
# #ok fuck it, we're manually calculating the margin of error
# #
# point_estimate = m$coef["miles", "(Estimate)"]
# se = m$coefficients["miles", "Std. Error"]
# tstar = qt(p = 0.975, df = m$df[2])
# moe = tstar*se
# point_estimate - moe
# point_estimate + moe
```

```{r}
point_estimate = coef(m)
```


### Morning Temperature \~ Impervious Surface + Distance to Coastline

```{r}
af_df_dists |>
  ggplot() +
  aes(x = miles, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles, data = am_df_dists) |>
  # compute the p-value, etc.
  summary()
m
```

### Afternoon Temperature \~ Impervious Surface + Distance to Coastline

```{r}
af_df_dists |>
  ggplot() +
  aes(x = miles, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles, data = af_df_dists) |>
  # compute the p-value, etc.
  summary()
m
```

### Evening Temperature \~ Impervious Surface + Distance to Coastline

```{r}
pm_df_dists |>
  ggplot() +
  aes(x = miles, y = temperature) +
  geom_smooth(method = "lm") + # line of best fit
  geom_point() # scatterplot

#lm for linear model

# do a linear regression
m = lm(temperature ~ percent_im + miles, data = pm_df_dists) |>
  # compute the p-value, etc.
  summary()
m
```
# Polynomial Model (using impervious surface in LOBF) - modeling air temperature versus distance to coastline 

```{r}
#wait a fucking minute... we include percent_im in this model so like this isn't just a clean parabolic relationship 
m = lm(
  temperature ~ 
    percent_im +
    time +
    time:miles +
    time:milessq,
  data = gen_temp
)
m %>% summary()

coef = (m %>% summary())$coefficient[, "Estimate"]

str_c("temperature = ", str_c(names(coef), " * ", coef, collapse = " + "))

coef
```


```{r}
#finding x and y for vertex -> aka for however many miles we move from the coastline, we go up y amount of degrees from intercept 

#am
am_vertex_x = -coef[["time6-7am:miles"]] / (2*coef[["time6-7am:milessq"]])
am_vertex_x

am_vertex_y = - (coef[["time6-7am:miles"]])^2 / (4*coef[["time6-7am:milessq"]])
am_vertex_y

am_average_at_coast = coef[["(Intercept)"]] + coef[["time6-7am"]]

#af
af_vertex_x = -coef[["time3-4pm:miles"]] / (2*coef[["time3-4pm:milessq"]])
af_vertex_x

af_vertex_y = - (coef[["time3-4pm:miles"]])^2 / (4*coef[["time3-4pm:milessq"]])
af_vertex_y

af_average_at_coast = coef[["(Intercept)"]]

#pm
pm_vertex_x = -coef[["time7-8pm:miles"]] / (2*coef[["time7-8pm:milessq"]])
pm_vertex_x

af_vertex_y = - (coef[["time7-8pm:miles"]])^2 / (4*coef[["time7-8pm:milessq"]])
af_vertex_y

af_average_at_coast = coef[["(Intercept)"]] + coef[["time7-8pm"]]
```

```{r}
gen_temp %>% 
  mutate(prediction = predict(m),
         residual = resid(m)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_0_percent_im = a*milessq + b*miles + c
  ) %>% 
  ggplot(aes(x = miles, y = temperature)) +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_0_percent_im), colour = "black") +
  facet_wrap(vars(time), scale = "free")
```

```{r}
gen_temp %>% 
  mutate(prediction = predict(m),
         residual = resid(m)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_avg_percent_im = a*milessq + b*miles + c + coef[["percent_im"]]*mean(gen_temp %>% 
  filter(time == "6-7am") %>% 
     pull(percent_im))
  ) %>% 
  ggplot(aes(x = miles, y = temperature)) +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_avg_percent_im), colour = "black") +
  labs(x = "Distance to Coastline (miles)",
       y =  "Air Temperature (°F)",
       title = "Air Temperature (°F) vs. Distance to Coastline (miles)") +
  facet_wrap(vars(time), scale = "free")
```

```{r}
gen_temp %>% 
  mutate(prediction = predict(m),
         residual = resid(m)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_0_percent_im = a*milessq + b*miles + c
  ) %>% 
filter(time == "6-7am") %>% 
  ggplot(aes(x = miles, y = temperature))  +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_0_percent_im), colour = "black") +
  coord_fixed(1/2)
```

```{r}
gen_temp %>% 
  mutate(prediction = predict(m),
         residual = resid(m)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_0_percent_im = a*milessq + b*miles + c
  ) %>% 
filter(time == "3-4pm") %>% 
  ggplot(aes(x = miles, y = temperature))  +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_0_percent_im), colour = "black") +
  #facet_wrap(vars(time), scale = "free") +
  coord_fixed(1/2)
```

```{r}
gen_temp %>% 
  mutate(prediction = predict(m),
         residual = resid(m)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_0_percent_im = a*milessq + b*miles + c
  ) %>% 
filter(time == "7-8pm") %>% 
  ggplot(aes(x = miles, y = temperature))  +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_0_percent_im), colour = "black") +
 # facet_wrap(vars(time), scale = "free") +
  coord_fixed(1/2)
```

# Polynomial Model (NOT using impervious surface in LOBF) - modeling air temperature versus distance to coastline 

```{r}
#wait a fucking minute... we include percent_im in this model so like this isn't just a clean parabolic relationship 
f = lm(
  temperature ~
    time +
    time:miles +
    time:milessq,
  data = gen_temp
)
f %>% summary()

coef = (f %>% summary())$coefficient[, "Estimate"]

str_c("temperature = ", str_c(names(coef), " * ", coef, collapse = " + "))

coef
```


```{r}
gen_temp %>% 
  mutate(prediction = predict(f),
         residual = resid(f)) %>% 
  group_by(time) %>% 
  mutate(
    a = case_when(
      time == "6-7am" ~ coef[["time6-7am:milessq"]],
      time == "3-4pm" ~ coef[["time3-4pm:milessq"]],
      time == "7-8pm" ~ coef[["time7-8pm:milessq"]]
    ),
    b = case_when(
      time == "6-7am" ~ coef[["time6-7am:miles"]],
      time == "3-4pm" ~ coef[["time3-4pm:miles"]],
      time == "7-8pm" ~ coef[["time7-8pm:miles"]]
    ),
    c = case_when(
      time == "6-7am" ~ coef[["(Intercept)"]] + coef[["time6-7am"]],
      time == "3-4pm" ~ coef[["(Intercept)"]],
      time == "7-8pm" ~ coef[["(Intercept)"]] + coef[["time7-8pm"]]
    ),
    prediction_for_avg_percent_im = a*milessq + b*miles + c) %>% 
  ggplot(aes(x = miles, y = temperature)) +
  geom_point(colour = "red") +
  geom_line(aes(y = prediction_for_avg_percent_im), colour = "black") +
  labs(x = "Distance to Coastline (miles)",
       y =  "Air Temperature (°F)",
       title = "Air Temperature (°F) vs. Distance to Coastline (miles)") +
  facet_wrap(vars(time), scale = "free")
```
